---
title: "PA1_template"
author: "danielMontjoy"
date: "Thursday, August 14, 2014"
output: html_document
---
###Loading the data
```{r load data}
library(ggplot2)
library(plyr)

if (!file.exists("activity.zip")) {
        download.file(url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",
                      destfile = "activity.zip",
                      method = "curl")
        }
if (!file.exists("activity.csv")) {
        unzip ("activity.zip")
        }


data<-read.csv("activity.csv");
mean(data$steps[which(data$steps>=0)])
```

### Cast to date

```{r fecha ,echo=TRUE}
data$date_final<- as.Date(data$date,"%Y-%m-%d")
```


###Clean Data
```{r ,echo=TRUE}
data2<-data[is.na(data$steps)==FALSE,]
```

### Histogram of the total number of steps taken each day  
```{r total_steps,echo=TRUE,fig.height=6}
pasos_anio<-aggregate(data2$steps,by=list(data2$date),sum,na.rm=TRUE)
names(pasos_anio)<-c("fecha","pasos")
hist(pasos_anio$pasos,xlab="Total Steps", main="Total Steps Histogram")
```
###Mean and Median  

```{r mean_median,echo=TRUE}
data_mean<-mean(data$steps[is.na(data$steps)==FALSE])

data_median<-median(data$steps[is.na(data$steps)==FALSE])

```

The mean is `r data_mean` and the median is `r data_median`  

###Series plot of the 5-minute interval  
```{r average_daily,echo=TRUE,fig.height=6}
pasos_interval<-aggregate(data2$steps,by=list(data2$interval),mean,na.rm=TRUE)
names(pasos_interval)<-c("intervalos","pasos")
plot(pasos_interval$pasos ~ pasos_interval$intervalos,type="l",xlab="Interval",ylab="Average Steps", main="Average Daily Activity Pattern")
```

```{r most,echo=TRUE}
most <- pasos_interval$intervalos[which.max(pasos_interval$pasos)]
```

The maximum number of steps in a 5 minute interval was taken on `r most`.

###Calculate and report the total number of missing values in the dataset  


```{r MissingValues,echo=TRUE}
mv<-colSums(is.na(data))
```

The total number of missing values is `r mv`.  

## The strategy for filling in all of the missing values in the dataset is put the mean  

```{r fill_In_MissingValues,echo=TRUE}

library(Hmisc)
data.imputed <- data
data.imputed$steps <- with(data.imputed, impute(steps, mean))

pasos_anio.imputed<-aggregate(data.imputed$steps,by=list(data.imputed$date),sum,na.rm=TRUE)
names(pasos_anio.imputed)<-c("fecha","pasos")
data_mean_imputed<-mean(pasos_anio.imputed$pasos)
data_median_imputed<-median(pasos_anio.imputed$pasos)


````
The mean and median total number of steps taken per day are `r data_mean_imputed` and `r data_median_imputed` respectively.**Imputing the missing data has increased the average number of steps.**
###Make a histogram of the total number of steps taken each day
```{r Hist2,echo=TRUE,fig.height=6}
hist(pasos_anio.imputed$pasos,xlab="Total Steps", main="Total Steps Histogram")
```

###Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r weekdays_weekends,echo=TRUE}
data.imputed$dia_semana<-weekdays(data.imputed$date_final)


data.imputed[,"day_factor"] <- ifelse(data.imputed$dia_semana == "sunday" |data.imputed$dia_semana =="saturday","Weekend", "Weekday")




```

###Make a panel plot containing a time series

```{r dayType, echo=TRUE, fig.height=6}
final <- aggregate(data.imputed$steps,
                   by=list(data.imputed$interval,
                           data.imputed$day_factor),
                   FUN=mean)
names(final) <- c("interval", "dayType", "steps")
xyplot(steps ~ interval | dayType, data = final, type = "l", layout=c(1,2))
```



